For finitary induction-Induction, Induction is Enough

### 2020-04-08 paper/paper.tex
some fixes suggested by Ambrus and the reviewers

\title{For Induction-Induction, Induction is Enough} %TODO Please add

@@

Many mutual inductive types can be reduced to indexed inductive types, where the
index disambiguates different sorts. An example is the predicates
$\mathsf{isEven}$ and $\mathsf{isOdd}$ defined mutually by the following
constructors.
\begin{alignat*}{5}
  & \mathsf{isEven} && : \N\ra\Set \\
  & \mathsf{isOdd} && : \N\ra \Set \\
  & \mathsf{zeroEven} && : \mathsf{isEven}\,\mathsf{zero} \\
  & \mathsf{sucEven} && : (n:\N)\ra\mathsf{isOdd}\,n\ra\mathsf{isEven}\,n \\
  & \mathsf{sucOdd} && : (n:\N)\ra\mathsf{isEven}\,n\ra\mathsf{isOdd}\,n
\end{alignat*}
These can be reduced to the following single inductive family.
\begin{alignat*}{5}
  & \mathsf{isEven?} && : \mathsf{Bool}\ra\N\ra\Set \\
  & \mathsf{zeroEven} && : \mathsf{isEven?}\,\mathsf{true}\,\mathsf{zero} \\
  & \mathsf{sucEven} && : (n:\N)\ra\mathsf{isEven?}\,\mathsf{false}\,n\ra\mathsf{isEven?}\,\mathsf{true}\,n \\
  & \mathsf{sucOdd} && : (n:\N)\ra\mathsf{isEven?}\,\mathsf{true}\,n\ra\mathsf{isEven?}\,\mathsf{false}\,n
\end{alignat*}

Inductive-inductive types (IITs \cite{forsberg-phd}) allow the mutual
definition of a type and a family of types over the first one. IITs
were originally introduced to represent the well-typed syntax of type
theory itself, and the main example is still the Chapman-style
\cite{chapman09eatitself} syntax for a type theory. A minimal use case
is the IIT of contexts and types given by the following constructors.
\begin{alignat*}{5}
  & \Con && : \Set \\
  & \Ty && : \Con\ra\Set \\
  & \mathsf{empty} && : \Con \\
  & \mathsf{ext} && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Con \\
  & \mathsf{U} && : (\Gamma:\Con)\ra\Ty\,\Gamma \\
  & \mathsf{El} && : (\Gamma:\Con)\ra\Ty\,(\mathsf{ext}\,\Gamma\,(\mathsf{U}\,\Gamma))
%  & \mathsf{Pi} && : (\Gamma:\Con)(A:\Ty\,\Gamma)\ra\Ty\,(\mathsf{ext}\,\Gamma\,A)\ra\,\Ty\,\Gamma
\end{alignat*}
This type has two sorts, $\Con$ and $\Ty$. The $\mathsf{ext}$ constructor of
$\Con$ refers to $\Ty$ and the $\Ty$-constructor $\mathsf{U}$ refers to $\Con$,
hence the two sorts have to be defined at the same time. Moreover, $\Ty$ is
indexed over $\Con$. This precludes a reduction analogous to the reduction of
$\mathsf{isEven}$--$\mathsf{isOdd}$, as we would get a type indexed over
itself. Another unique feature of IITs (which also shows up in higher inductive
types \cite{HoTTbook}) is that later constructors can refer to previous
constructors: in our case, $\mathsf{El}$ mentions
$\mathsf{ext}$.
% \[
%   \mathsf{Con?} : (b:\mathsf{Bool})\ra\mathsf{if}\,b\,\mathsf{then}\,\Set\,\mathsf{else}\,(\mathsf{Con?}\,\mathsf{false}\ra\Set),
% \]

The elimination principle for the above IIT has two motives and one method
for each constructor. It is an example of a \emph{recursive-recursive}
function (following the nomenclature of \cite{forsberg-phd}) which
means that there are two mutually defined functions, where the type of
the second function depends on the first function. The proof assistant
Agda \cite{norell07thesis} allows the definition of such functions
(even from non-IITs) and it is also currently the only proof assistant
supporting IITs\footnote{An experimental version of Coq with IITs is also
  available on github.}. The elimination principle for $\Con$--$\Ty$ says that
given the following components,
\begin{alignat*}{5}
  & \Con^D && : \Con\ra\Set \\
  & \Ty^D && : \Con^D\,\Gamma\ra\Ty\,\Gamma\ra\Set \\
  & \mathsf{empty}^D && : \Con^D\,\mathsf{empty} \\
  & \mathsf{ext}^D && : (\Gamma^D:\Con^D\,\Gamma)\ra \Ty^D\,\Gamma^D\,A\ra \Con^D\,(\mathsf{ext}\,\Gamma\,A) \\
  & \U^D && : (\Gamma^D:\Con^D\,\Gamma)\ra \Ty^D\,\Gamma^D\,(\mathsf{U}\,\Gamma) \\
  & \El^D && : (\Gamma^D:\Con^D\,\Gamma)\ra\,\Ty^D\,(\mathsf{ext}^D\,\Gamma^D\,(\U^D\,\Gamma^D))\,(\mathsf{El}\,\Gamma)
%  & Pi^D && : (\Gamma^D:\Con^D\,\Gamma)(A^D:\Ty^D\,\Gamma^D\,A)\ra \Ty^D\,(\mathsf{ext}^D\,\Gamma^D\,A^D)\ra\,\Ty^D\,\Gamma^D\,(\mathsf{Pi}\,\Gamma\,A\,B)
\end{alignat*}
we obtain two functions
\begin{alignat*}{5}
  & \mathsf{elimCon} && : (\Gamma:\Con)\ra \Con^D\,\Gamma \\
  & \mathsf{elimTy} && : (A:\Ty\,\Gamma)\ra \Ty^D\,(\mathsf{elimCon}\,\Gamma)
\end{alignat*}
with the following computation rules.
\begin{alignat*}{5}
  & \mathsf{elimCon}\,\mathsf{empty} && = \mathsf{empty}^D \\
  & \mathsf{elimCon}\,(\mathsf{ext}\,\Gamma\,A) && = \mathsf{ext}^D\,(\mathsf{elimCon}\,\Gamma)\,(\mathsf{elimTy}\,A) \\
  & \mathsf{elimTy}\,(\mathsf{U}\,\Gamma) && = \U^D\,(\mathsf{elimCon}\,\Gamma) \\
  & \mathsf{elimTy}\,(\mathsf{El}\,\Gamma) && = \El^D\,(\mathsf{elimCon}\,\Gamma)
\end{alignat*}

@@

Note that initiality refers to a notion of homomorphism between models. As
of now, there is no general metatheory for (large, infinitary) QIITs from which
this notion would fall out. Hence, we give an ad-hoc definition for
homomorphisms, which can be summarized as a functor between underlying
categories which strictly preserves all additional family structure and type
formers.

@@

Natural numbers:
\[
\boldsymbol{\cdot}\,\ext\,N:\U\,\ext\,z : \El\,N\,\ext\,s : N \Ra \El\,N
\]
Lists with elements of a given $A : \Set$ type. Here we use the $\Pim$ function
to include an external type into the signature.
\[
\boldsymbol{\cdot}\,\ext\,L:\U\,\ext\,nil : \El\,L\,\ext\,
      cons : \Pim\,A\,(\lambda \_.\, L\Ra L)
\]
The inductive-inductive example from Section \ref{sec:intro}:
\begin{alignat*}{3}
  & \boldsymbol{\cdot} \,\ext\,Con : \U\,\ext\,Ty : Con\Ra\U\,\ext\,empty : \El\,Con\,\ext\,ext : \Pi\,(\Gamma : Con)\,(Ty\oldapp\Gamma \Ra Con) \\
  & \hspace{0.5em}\,\ext\, U : \Pi\,(\Gamma:Con)\,(\El\,(Ty\oldapp\Gamma))
    \,\ext\,El : \Pi\,(\Gamma : Con)(\El\,(Ty\oldapp(ext\oldapp\Gamma\oldapp (U\oldapp\Gamma))))
\end{alignat*}

>>>>

\title{For finitary induction-Induction, Induction is Enough} %TODO Please add

@@

Many mutual inductive types can be reduced to indexed inductive types, where the
index disambiguates different sorts. An example is the predicates (or \emph{indexed sorts})
$\mathsf{isEven}$ and $\mathsf{isOdd}$ defined mutually by the following
constructors.
\begin{alignat*}{5}
  & \mathsf{isEven} && : \N\ra\Set \\
  & \mathsf{isOdd} && : \N\ra \Set \\
  & \mathsf{zeroEven} && : \mathsf{isEven}\,\mathsf{zero} \\
  & \mathsf{sucEven} && : (n:\N)\ra\mathsf{isOdd}\,n\ra\mathsf{isEven}\,(n+1) \\
  & \mathsf{sucOdd} && : (n:\N)\ra\mathsf{isEven}\,n\ra\mathsf{isOdd}\,(n+1)
\end{alignat*}
These can be reduced to the following single inductive family.
\begin{alignat*}{5}
  & \mathsf{isEven?} && : \mathsf{Bool}\ra\N\ra\Set \\
  & \mathsf{zeroEven} && : \mathsf{isEven?}\,\mathsf{true}\,\mathsf{zero} \\
  & \mathsf{sucEven} && : (n:\N)\ra\mathsf{isEven?}\,\mathsf{false}\,n\ra\mathsf{isEven?}\,\mathsf{true}\,(n+1) \\
  & \mathsf{sucOdd} && : (n:\N)\ra\mathsf{isEven?}\,\mathsf{true}\,n\ra\mathsf{isEven?}\,\mathsf{false}\,(n+1)
\end{alignat*}

Inductive-inductive types (IITs \cite{forsberg-phd}) allow the mutual
definition of a type and a family of types over the first one. IITs
were originally introduced to represent the well-typed syntax of type
theory itself, and the main example is still the Chapman-style
\cite{chapman09eatitself} syntax for a type theory. A minimal use case
is the IIT of contexts and types given by the following constructors.
\begin{alignat*}{5}
  & \Con && : \Set \\
  & \Ty && : \Con\ra\Set \\
  & \mathsf{empty} && : \Con \\
  & \mathsf{ext} && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Con \\
  & \mathsf{U} && : (\Gamma:\Con)\ra\Ty\,\Gamma \\
  & \mathsf{El} && : (\Gamma:\Con)\ra\Ty\,(\mathsf{ext}\,\Gamma\,(\mathsf{U}\,\Gamma))
%  & \mathsf{Pi} && : (\Gamma:\Con)(A:\Ty\,\Gamma)\ra\Ty\,(\mathsf{ext}\,\Gamma\,A)\ra\,\Ty\,\Gamma
\end{alignat*}
This type has two sorts, $\Con$ and $\Ty$. The $\mathsf{ext}$ constructor of
$\Con$ refers to $\Ty$ and the $\Ty$-constructor $\mathsf{U}$ refers to $\Con$,
hence the two sorts have to be defined simultaneously. Moreover, $\Ty$ is
indexed over $\Con$. This precludes a reduction analogous to the reduction of
$\mathsf{isEven}$--$\mathsf{isOdd}$, as we would get a type indexed over
itself. Another unique feature of IITs (which also shows up in higher inductive
types \cite{HoTTbook}) is that later constructors can refer to previous
constructors: in our case, $\mathsf{El}$ mentions
$\mathsf{ext}$.
% \[
%   \mathsf{Con?} : (b:\mathsf{Bool})\ra\mathsf{if}\,b\,\mathsf{then}\,\Set\,\mathsf{else}\,(\mathsf{Con?}\,\mathsf{false}\ra\Set),
% \]

The elimination principle for the above IIT has two motives and one method
for each constructor. It is an example of a \emph{recursive-recursive}
function (following the nomenclature of \cite{forsberg-phd}) which
means that there are two mutually defined functions, where the type of
the second function depends on the first function. The proof assistant
Agda \cite{norell07thesis} allows the definition of such functions
(even from non-IITs) and it is also currently the only proof assistant
supporting IITs\footnote{An experimental version of Coq with IITs is also
  available on github.}. The elimination principle for $\Con$--$\Ty$ says that
given the following components,
where the variable $\Gamma$ is implicitly universally quantified when required
(we indeed sometimes omit universal quantification in the following, for readability),
\begin{alignat*}{5}
  & \Con^D && : \Con\ra\Set \\
  & \Ty^D && : \Con^D\,\Gamma\ra\Ty\,\Gamma\ra\Set \\
  & \mathsf{empty}^D && : \Con^D\,\mathsf{empty} \\
  & \mathsf{ext}^D && : (\Gamma^D:\Con^D\,\Gamma)\ra \Ty^D\,\Gamma^D\,A\ra \Con^D\,(\mathsf{ext}\,\Gamma\,A) \\
  & \U^D && : (\Gamma^D:\Con^D\,\Gamma)\ra \Ty^D\,\Gamma^D\,(\mathsf{U}\,\Gamma) \\
  & \El^D && : (\Gamma^D:\Con^D\,\Gamma)\ra\,\Ty^D\,(\mathsf{ext}^D\,\Gamma^D\,(\U^D\,\Gamma^D))\,(\mathsf{El}\,\Gamma)
%  & Pi^D && : (\Gamma^D:\Con^D\,\Gamma)(A^D:\Ty^D\,\Gamma^D\,A)\ra \Ty^D\,(\mathsf{ext}^D\,\Gamma^D\,A^D)\ra\,\Ty^D\,\Gamma^D\,(\mathsf{Pi}\,\Gamma\,A\,B)
\end{alignat*}
we obtain two functions
\begin{alignat*}{5}
  & \mathsf{elimCon} && : (\Gamma:\Con)\ra \Con^D\,\Gamma \\
  & \mathsf{elimTy} && : (A:\Ty\,\Gamma)\ra \Ty^D\,(\mathsf{elimCon}\,\Gamma)
\end{alignat*}
with the following computation rules.
\begin{alignat*}{5}
  & \mathsf{elimCon}\,\mathsf{empty} && = \mathsf{empty}^D \\
  & \mathsf{elimCon}\,(\mathsf{ext}\,\Gamma\,A) && = \mathsf{ext}^D\,(\mathsf{elimCon}\,\Gamma)\,(\mathsf{elimTy}\,A) \\
  & \mathsf{elimTy}\,(\mathsf{U}\,\Gamma) && = \U^D\,(\mathsf{elimCon}\,\Gamma) \\
  & \mathsf{elimTy}\,(\mathsf{El}\,\Gamma) && = \El^D\,(\mathsf{elimCon}\,\Gamma)
\end{alignat*}

@@

Note that initiality refers to a notion of homomorphism between models. As
of now, there is no general metatheory for (large, infinitary) QIITs from which
this notion would be deduced. Hence, we give an ad-hoc definition for
homomorphisms, which can be summarized as a functor between underlying
categories which strictly preserves all additional family structure and type
formers.

@@

Natural numbers:
\[
\boldsymbol{\cdot}\,\ext\,N:\U\,\ext\,z : \El\,N\,\ext\,s : N \Ra \El\,N
\]
Lists with elements of a given $A : \Set$ type. Here we use the $\Pim$ function
to include an external type into the signature.
\[
\boldsymbol{\cdot}\,\ext\,L:\U\,\ext\,nil : \El\,L\,\ext\,
      cons : \Pim\,A\,(\lambda \_.\, L\Ra \El\,L)
\]
The inductive-inductive example from Section \ref{sec:intro}:
\begin{alignat*}{3}
  & \boldsymbol{\cdot} \,\ext\,Con : \U\,\ext\,Ty : Con\Ra\U\,\ext\,empty : \El\,Con\,\ext\,ext : \Pi\,(\Gamma : Con)\,(Ty\oldapp\Gamma \Ra Con) \\
  & \hspace{0.5em}\,\ext\, U : \Pi\,(\Gamma:Con)\,(\El\,(Ty\oldapp\Gamma))
    \,\ext\,El : \Pi\,(\Gamma : Con)(\El\,(Ty\oldapp(ext\oldapp\Gamma\oldapp (U\oldapp\Gamma))))
\end{alignat*}

### 2020-04-08 paper/universaliit.tex
some fixes suggested by Ambrus and the reviewers

\subsubsection{Untyped syntax}
The untyped syntax is defined as the following inductive datatype. Variables are
modeled as de Bruijn indices, i.e.\ as natural numbers pointing to a position in
the context. We use the additional default constructor $\erru:\Tmu$ in case of
error (ill-scoped substitution). The typing judgments will not mention $\erru$.
\begin{alignat*}{5}
  & \rlap{(1) Substitution calculus} \\[0.5em]
  & \Conu && : \Set \\
  & \Tyu  && : \Set \\
  & \Subu  && : \Set \\
  % & \Subpre  && : \Set \\
  & \Tmu  && : \Set \\
  & \emptyu && : \Conu \\
    & \epsilonu && : \Subu \\
  & \blank\extu\blank && : \Conu\ra\Tyu\ra\Conu \\
  & \blank\consu\blank && : \Subu\ra\Tmu\ra\Subu \\
  & \varu  && : \N \ra \Tmu \\
  & \rlap{(2) Universe} \\[0.5em]
  & \Uu && : \Tyu \\
  & \Elu && : \Tmu\ra\Tyu \\
  & \rlap{(3) Inductive parameters} \\[0.5em]
  & \Piu && : \Tmu\ra\Tyu\ra\Tyu \\
  & \blank\appu\blank && : \Tmu\ra\Tmu\ra \Tmu \\
  & \rlap{(4-5) Metatheoretic parameters} \\[0.5em]
  & \Pimu && : (T:\Set)\ra(T\ra\Tyu)\ra\Tyu \\
  & \Piiu && : (T:\Set)\ra(T\ra\Tmu)\ra\Tmu \\
  & \blank\appmu \blank && : \Tmu\ra(\alpha:T) \ra\Tmu\\
  & \rlap{(6) Default value} \\[0.5em]
  & \erru && : \Tmu
\end{alignat*}

@@

  \subsubsection{Exchange laws: weakening and substitution}
  Many lemmas for types and terms are shown by induction on the untyped syntax.
  Below, $Z$ denotes either a term or a type.
\begin{alignat*}{5}
  % liftV (S (n + p)) (liftV n q) ≡ liftV n (liftV (n + p) q)
  & \wk\mhyphen\wk && : \wk_{n+p+1} (\wk_n\,Z) = \wk_n(\wk_{n+p}\,Z)\\
  % n-subZ-wkZ : ∀ n A z → (lifZZ n A) [ n ↦ z ]Z  ≡ A
  & \wk_n[n] && : (\wk_{n}\,Z)[ n := z] = Z \\
  % lifZ-l-subV : ∀ n p u x → lifZZ (n + p) (x [ n ↦ u ]V) ≡ (lifZV (S (n + p)) x) [ n ↦ (lifZZ p u) ]V
  & \wk_+[] && : (\wk_{n+p+1}\,Z)[ n := \wk_p\,u] = \wk_{n+p}\,(Z[n := u]) \\
  % l-subV-lifZV : ∀ Δ u n x → (lifZV n x) [ (S (n + Δ)) ↦ u ]V  ≡ lifZZ n (x [ (n + Δ) ↦ u ]V)
  & \wk[+] && : (\wk_{n}\,Z)[ n+p+1 :=  u] = \wk_{n}\,(Z[n+p := u]) \\
  % l-subV-l-subV : ∀ n p z u x →  (x [ n ↦ u ]V) [ (n + p) ↦ z ]Z  ≡ (x [ (S (n + p)) ↦ z ]V) [ n ↦ (u [ p ↦ z ]Z) ]Z
  & [][+] && : Z[n :=u][n+p:=z] = Z[n+p+1:=z][n:= (u[p:=z])]
  \end{alignat*}
  Below, $\keep^n$ denotes $\keep$ iterated $n$ times.
\begin{alignat*}{5}
  % liftV (S (n + p)) (liftV n q) ≡ liftV n (liftV (n + p) q)
  % liftV=wkS
  & [\keep^n\mhyphen \wkO] && : Z[\keep^n\,(\wkO\,\sigma)] = \wk_n (Z[\keep^n\,\sigma])
  \\
  % lifZₛV : ∀ n xp σp Zp → (lifZV n xp [ iZer n keep (Zp ∷ σp) ]V) ≡ (xp [ iZer n keep σp ]V)
  & \wk_n[\keep^n\mhyphen\cons]  & &
  :
  (\wk_n\,Z) [\keep^n\, (u \consu \sigma)] = Z [ \keep^n \sigma]
  \\
  % l-sub[]V : ∀ n x z σ
    % ((x [ n := z ]V) [ iZer n keep σ ]Z) = (x [ iZer (S n) keep σ ]V) [ n := (z [ σ ]Z) ]Z
   & [:=][\keep] && : Z [ n:= u][\keep^n\,\sigma] =  Z[\keep^{n+1}\,\sigma][n := u[\sigma]]
  \end{alignat*}
As particular cases for $n=0$, we get
\begin{alignat*}{5}
  & {}{\circ}{}\wkO & & : \sigma \circ (\wkO \tau) = \wkO (\sigma \circ \tau) \\
  & \wkO{}{\circ}{}, && : \wkO\,\sigma \circ (t \consu \tau ) = \sigma \circ \tau \\
  & [\wkO] && : t[\wkO\,\sigma] = \wk_0 (t[\sigma]) \\
  & \wk_0[\cons] && : (\wk_0 Z)[u \consu \sigma] = Z[\sigma] \\
  & [0:=][] && : Z [ 0:= u][ \sigma] =  Z[\keep\,\sigma][0 := u[\sigma]]
\end{alignat*}
Finally, we show the following:
\begin{alignat*}{5}
  & [][] && : Z[\sigma][\tau] = Z[\sigma\circ\tau]
  \\
  & \ass && : (\sigma\circ \delta)\circ\tau = \sigma\circ (\delta\circ\tau)
  \end{alignat*}
We defer laws for identity substitutions after the definition of the typing
judgments, as the proofs require that some inputs are well-typed.

@@

This approach is an alternative version of Streicher's method for interpreting
preterms in an arbitrary model \cite{streichersemantics}. Streicher first defines
a family of partial maps from the presyntax to a model, then shows that the maps
are total on well-formed input. We have found that our approach is significantly
easier to formalise. Too see why, note that the right notion of partial map in
type theory, which does not presume decidable definedness, is fairly
heavyweight:
\[
  \mathsf{PartialMap}\,A\,B := A\ra \big((P : \Set)\times\isaprop\,P\times (P \ra B)\big)
\]
In the above definition, we notice an opportunity for converting a fibered
definition of a type family into an indexed one; if we drop the propositionality
for $P$ for the time being, we may equivalently return a family indexed over
$B$, which is exactly just a relation $A\ra B\ra\Set$. Then, in our approach, we
recover uniqueness of $P$ through the functionality requirement on the $A\ra
B\ra\Set$ relation, and totality by already assuming well-formedness of $A$. In
type theory, using indexed families instead of display maps is a common
convenience, since the former are natively supported, while the
latter require carrying around auxiliary propositional equalities.

@@

  UIP \cite{hugunin2019constructing} in cubical type theory, and demonstrated

>>>>

\subsubsection{Untyped syntax}
The untyped syntax is defined as the following inductive datatype. Variables are
modeled as de Bruijn indices, i.e.\ as natural numbers pointing to a position in
the context. We use the additional default constructor $\erru:\Tmu$ in case of
error (ill-scoped substitution). The typing judgments will not mention $\erru$.
The main interest of $\erru$ is that it behaves like a closed term (which the
theory of signatures lacks), in the sense that it is
invariant under substitution. This makes expected equalities about 
substitution (see Section~\ref{ss:exch-laws}) true even in the ill-typed case, thus reducing the amount of hypotheses.
\begin{alignat*}{5}
  & \rlap{(1) Substitution calculus} \\[0.5em]
  & \Conu && : \Set \\
  & \Tyu  && : \Set \\
  & \Subu  && : \Set \\
  % & \Subpre  && : \Set \\
  & \Tmu  && : \Set \\
  & \emptyu && : \Conu \\
    & \epsilonu && : \Subu \\
  & \blank\extu\blank && : \Conu\ra\Tyu\ra\Conu \\
  & \blank\consu\blank && : \Subu\ra\Tmu\ra\Subu \\
  & \varu  && : \N \ra \Tmu \\
  & \rlap{(2) Universe} \\[0.5em]
  & \Uu && : \Tyu \\
  & \Elu && : \Tmu\ra\Tyu \\
  & \rlap{(3) Inductive parameters} \\[0.5em]
  & \Piu && : \Tmu\ra\Tyu\ra\Tyu \\
  & \blank\appu\blank && : \Tmu\ra\Tmu\ra \Tmu \\
  & \rlap{(4-5) Metatheoretic parameters} \\[0.5em]
  & \Pimu && : (T:\Set)\ra(T\ra\Tyu)\ra\Tyu \\
  & \Piiu && : (T:\Set)\ra(T\ra\Tmu)\ra\Tmu \\
  & \blank\appmu \blank && : \Tmu\ra(\alpha:T) \ra\Tmu\\
  & \rlap{(6) Default value} \\[0.5em]
  & \erru && : \Tmu
\end{alignat*}

@@

  \subsubsection{Exchange laws: weakening and substitution}
  \label{ss:exch-laws}
  Many lemmas for types and terms are shown by induction on the untyped syntax.
  Below, $Z$ denotes either a term or a type.
\begin{alignat*}{5}
  % liftV (S (n + p)) (liftV n q) ≡ liftV n (liftV (n + p) q)
  & \wk\mhyphen\wk && : \wk_{n+p+1} (\wk_n\,Z) = \wk_n(\wk_{n+p}\,Z)\\
  % n-subZ-wkZ : ∀ n A z → (lifZZ n A) [ n ↦ z ]Z  ≡ A
  & \wk_n[n] && : (\wk_{n}\,Z)[ n := z] = Z \\
  % lifZ-l-subV : ∀ n p u x → lifZZ (n + p) (x [ n ↦ u ]V) ≡ (lifZV (S (n + p)) x) [ n ↦ (lifZZ p u) ]V
  & \wk_+[] && : (\wk_{n+p+1}\,Z)[ n := \wk_p\,u] = \wk_{n+p}\,(Z[n := u]) \\
  % l-subV-lifZV : ∀ Δ u n x → (lifZV n x) [ (S (n + Δ)) ↦ u ]V  ≡ lifZZ n (x [ (n + Δ) ↦ u ]V)
  & \wk[+] && : (\wk_{n}\,Z)[ n+p+1 :=  u] = \wk_{n}\,(Z[n+p := u]) \\
  % l-subV-l-subV : ∀ n p z u x →  (x [ n ↦ u ]V) [ (n + p) ↦ z ]Z  ≡ (x [ (S (n + p)) ↦ z ]V) [ n ↦ (u [ p ↦ z ]Z) ]Z
  & [][+] && : Z[n :=u][n+p:=z] = Z[n+p+1:=z][n:= (u[p:=z])]
  \end{alignat*}
  Below, $\keep^n$ denotes $\keep$ iterated $n$ times.
\begin{alignat*}{5}
  % liftV (S (n + p)) (liftV n q) ≡ liftV n (liftV (n + p) q)
  % liftV=wkS
  & [\keep^n\mhyphen \wkO] && : Z[\keep^n\,(\wkO\,\sigma)] = \wk_n (Z[\keep^n\,\sigma])
  \\
  % lifZₛV : ∀ n xp σp Zp → (lifZV n xp [ iZer n keep (Zp ∷ σp) ]V) ≡ (xp [ iZer n keep σp ]V)
  & \wk_n[\keep^n\mhyphen\cons]  & &
  :
  (\wk_n\,Z) [\keep^n\, (u \consu \sigma)] = Z [ \keep^n \sigma]
  \\
  % l-sub[]V : ∀ n x z σ
    % ((x [ n := z ]V) [ iZer n keep σ ]Z) = (x [ iZer (S n) keep σ ]V) [ n := (z [ σ ]Z) ]Z
   & [:=][\keep] && : Z [ n:= u][\keep^n\,\sigma] =  Z[\keep^{n+1}\,\sigma][n := u[\sigma]]
  \end{alignat*}
As particular cases for $n=0$, we get
\begin{alignat*}{5}
  & {}{\circ}{}\wkO & & : \sigma \circ (\wkO \tau) = \wkO (\sigma \circ \tau) \\
  & \wkO{}{\circ}{}, && : \wkO\,\sigma \circ (t \consu \tau ) = \sigma \circ \tau \\
  & [\wkO] && : t[\wkO\,\sigma] = \wk_0 (t[\sigma]) \\
  & \wk_0[\cons] && : (\wk_0 Z)[u \consu \sigma] = Z[\sigma] \\
  & [0:=][] && : Z [ 0:= u][ \sigma] =  Z[\keep\,\sigma][0 := u[\sigma]]
\end{alignat*}
Finally, we show the following:
\begin{alignat*}{5}
  & [][] && : Z[\sigma][\tau] = Z[\sigma\circ\tau]
  \\
  & \ass && : (\sigma\circ \delta)\circ\tau = \sigma\circ (\delta\circ\tau)
  \end{alignat*}
We defer laws for identity substitutions after the definition of the typing
judgments, as the proofs require that some inputs are well-typed.

@@

This approach is an alternative version of Streicher's method for interpreting
preterms in an arbitrary model \cite{streichersemantics}. Streicher first defines
a family of partial maps from the presyntax to a model, then shows that the maps
are total on well-formed input. We have found that our approach is significantly
easier to formalise. To see why, note that the right notion of partial map in
type theory, which does not presume decidable definedness, is fairly
heavyweight:
\[
  \mathsf{PartialMap}\,A\,B := A\ra \big((P : \Set)\times\isaprop\,P\times (P \ra B)\big)
\]
In the above definition, we notice an opportunity for converting a fibered
definition of a type family into an indexed one; if we drop the propositionality
for $P$ for the time being, we may equivalently return a family indexed over
$B$, which is exactly just a relation $A\ra B\ra\Set$. Then, in our approach, we
recover uniqueness of $P$ through the functionality requirement on the $A\ra
B\ra\Set$ relation, and totality by already assuming well-formedness of $A$. In
type theory, using indexed families instead of display maps is a common
convenience, since the former are natively supported, while the
latter require carrying around auxiliary propositional equalities.

@@

  UIP \cite{jasper} in cubical type theory, and demonstrated
